---
title: "Assignment 2: Data Wrangling with Base R"
author: "Louis Dion"
date: "September 12, 2019"
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

# Questions
## 1. Download the c2015 dataset to your computer. Use function getwd() to check the current working directory. Use setwd() to change the current directory to the c2015 file.
```{r}
getwd()
setwd("C:/Users/student/Documents/MATH421/data")
#Will not allow me to set the directory to the exact file 
#additionally it only sets the directory for this chunk and resets for other chunks
```

## 2. We need to install a package to read the xlsx file. (Let’s not change the xlsx to csv here) There are a few packages for this. I recommend to use the readxl package. This package is contained in the tidyverse package so if you already installed tidyverse, you should have it already. If not, install and load the readxl package
```{r}
library(readxl)
```

## 3. Use read_excel() to read the c2015 dataset. Use function class() to check the type of data you just read in. You will notice that the data now is not just a data frame, it is also a tibble. A tibble is a generalization of a data frame, so you can still use all the functions and syntax for data frame with tibble.
```{r}
c2015<-read_excel('MATH421/data/c2015.xlsx') 
#using directory that was reset to because directory did not carry over to this chunk
class(c2015)
```

## 4.Use dim function to check the dimension of the data. Since this data is quite big, a common practice is to randomly subset the data to analyze. Use sample function to create a new dataset that has a random 1000 observations from the original data. Use set.seed(2019) before using the sample function to set the seed for the randomness so that everyone in class is working with the same random subset of the data.
```{r}
dim(c2015)
set.seed(2019)
sample2015<-c2015[sample(nrow(c2015),1000),] #########
sample2015
```

## 5. Use summary function to have a quick look at the data. You will notice there is one variable is actually a constant. Remove that variable from the data.
```{r}
summary(sample2015)
sample2015<-subset(sample2015,select= -YEAR)
```

## 6. Check the number of missing values (NA) in each column.
```{r}
x<-is.na(sample2015)
sum(x)
colSums(x)
```

## 7. There are missing values in this data that are not NAs. Identify the form of these missing values. Check the number of these missing values in each column. Notice that you may want to use na.rm = TRUE when counting these missing values.
```{r}
sample2015
#By looking at the sample dataset, some variable have values "Unknown" "Not Rep" that are missing values, need to replace Unknown and Not Rep to NA

sample2015<-replace(sample2015,sample2015=="Unknown"|sample2015=="Not Rep",NA)
sample2015

x<-is.na(sample2015)
sum(x)
colSums(x)

```

## 8. Change the missing values in SEX variable to “Female”
```{r}
sample2015['SEX'][is.na(sample2015['SEX'])]<-"Female"
sample2015
```

## 9. Fix the AGE variable so that it is in the right form and has no missing values. 
```{r}
#step 1
sample2015['AGE'][sample2015['AGE']=='Less than 1']<-'0'
#step 2
sample2015$AGE<-as.numeric(sample2015$AGE)
#step 3
sample2015$AGE[is.na(sample2015$AGE)]<-mean(sample2015$AGE,na.rm=TRUE)
sample2015
```

## 10. Put the TRAV_SP(Travel Speed) variable in the right form (type) and remove all missing values. Calculate the average speed. You can use a non-base R function for this question.
```{r}
sample2015$TRAV_SP<-substr(sample2015$TRAV_SP, 1, nchar(sample2015$TRAV_SP)-4)
sample2015$TRAV_SP<-as.numeric(as.character(sample2015$TRAV_SP))
nomissing<-sample2015[!is.na(sample2015$TRAV_SP),]
mean(nomissing$TRAV_SP)
```

## 11. Compare the average speed of those who had "No Apprent Injury" and the rest. What do you observe?
```{r}
noinjury<-nomissing[nomissing$INJ_SEV=='No Apparent Injury (O)',]
injury<-nomissing[nomissing$INJ_SEV!='No Apparent Injury (O)',]
mean(noinjury$TRAV_SP,na.rm=TRUE)  
mean(injury$TRAV_SP,na.rm=TRUE)

# Travel speed with injury is higher than travel speed without injury
```

## 12. Use the SEAT_POS variable to filter the data so that there is only drivers in the dataset. Compare the average speed of man drivers and woman drivers. Comment on the results.
```{r}
driver<-nomissing[nomissing$SEAT_POS=="Front Seat, Left Side",]
maledriver<-driver[driver$SEX=='Male',]
femaledriver<-driver[driver$SEX=='Female',]
mean(maledriver$TRAV_SP,na.rm=TRUE)
mean(femaledriver$TRAV_SP,na.rm=TRUE)
#Males in accidents tend to drive faster than females in accidents
```

## 13.Compare the average speed of drivers who drink and those who do not. Comment on the results. 
```{r}
drink<-driver[driver$DRINKING=='Yes (Alcohol Involved)',]
nodrink<-driver[driver$DRINKING=='No (Alcohol Not Involved)',]
mean(drink$TRAV_SP,na.rm=TRUE)
mean(nodrink$TRAV_SP,na.rm=TRUE)

#People who drink tended to drive faster than people who did not drink
```

## 14. Hypothesize about the age range of drivers who may drive more aggressively. Test your hypothesis by comparing the average speed of those in this age range and the rest. Comment on the results.
```{r}
#I am hypothesizing that drivers under the age of 30 drive more aggressively, meaning they drive faster, than people aged 30 or above
young<-driver[driver$AGE<30,]
old<-driver[driver$AGE>=30,]
mean(young$TRAV_SP,na.rm=TRUE)
mean(old$TRAV_SP,na.rm=TRUE)
#These results show that people aged under 30 tended to drive faster than people aged 30 or older
```

## 15.  If the data did not confirm your hypothesis in 14. Could you identify an age group of drivers who may drive more aggressively?
```{r}
# Number 14 did confirm my hypothesis, so I do not need to complete this step.
```
