---
title: 'Assignment 1: Data Wrangling With Base R'
author: "Louis Dion"
date: "September 9, 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# Part 1 - R Basics
## 1. Calculate the following sums.

```{r}
x<-c(1:2019)
s1<-sum(x)
s1
s2<-sum(x^3)
s2
s3<-sum(x^x)
s3
y<-x*c(1,-1)
s4<-sum(y^x)
s4
s5<-sum((1/(x^2)))
s5
s6<-sum((1/x))
s6
s7<-sum((1/(x^3)))
s7
s8<-sum((1/y))
s8
```
## 2. Rnorm 1000 samples of mean 10 and sd 1
```{r}
x<-rnorm(1000,mean=10,sd=1)
```

### a. mean and standard deviation
```{r}
mean(x)
sd(x)
```
### b. How many do you think are greater than 10? Check estimation. 
I believe there are 500 samples that are greater than 10.
```{r}
y<-x>10
sum(y)
```
### c. Histogram of sample
```{r}
hist(x)
```
### .d Estimate P(x>1) for N(2,1)
```{r}
z<-rnorm(1000, mean=2, sd=1)
w<-z>1
sum(w)/1000
```
## 3. Tossing a fair dice
### a. Generate sample of 1000 values with replacment
```{r}
x<-sample(c(1:6),1000,replace=TRUE)
x
```
### b. mean and sd of sample
```{r}
mean(x)
sd(x)
```
### c. how many times 6 occurs
```{r}
y<-x==6
sum(y)
```
### d. Table function to show frequency of values
```{r}
table(x)
```

### e. Relative Frequency of values
```{r}
prop.table(table(x))
```
### f. Plot Frequency of values
```{r}
hist(x)
```
## 4. Experiment tossing 3 dice
```{r}
X1<-sample(c(1:6),1000,replace=TRUE)
X2<-sample(c(1:6),1000,replace=TRUE)
X3<-sample(c(1:6),1000,replace=TRUE)
```
### a. P(X1>X2+X3) 
```{r}
y<-X1>(X2+X3)
sum(y)/1000
```

### b. (P(X1^2>X2^2+X3^2))
```{r}
z<- X1^2>X2^2+X3^2
sum(z)
```
## 5. Estimate probability of 3 tails in a row (using matrix)
```{r}
x<- matrix(sample(c(0,1),3000,replace=TRUE), (nrows=1000))
y<-rowSums(x)==3
sum(y)/1000
```
## 6. Extra Credit

## 7. Central Limit Theorem
### Generate 100 samples of 1000 observations
```{r}
A<-matrix(sample(runif(n=10^5,min=0,max=1)),(nrows=100))
```
### Means of each sample
```{r}
x<-rowSums(A)/1000
```
### Plot histogram of x
```{r}
hist(x)
```
#### yes, the distribution looks normal
### Up the numbers (500 samples of 2000 observations)
```{r}
A<-matrix(sample(runif(n=10^6,min=0,max=1)),(nrows=500))
x<-rowSums(A)/1000
hist(x)
```
#### This distribution looks more normal
### Two more distributions 
#### Exponential Distribution (n=1000, lambda rate=.2) per sample
```{r}
A<-matrix(sample(rexp(n=10^5,rate=.2)),(nrows=100))
x<-rowSums(A)/1000
hist(x)
```
##### histogram looks relatively normal
#### Poisson distribution (n=1000, lambda=.2) per sample
```{r}
A<-matrix(sample(rpois(n=10^5,lambda=.2)),(nrows=100))
x<-rowSums(A)/1000
hist(x)
```
##### histogram looks normal
# Part 2 - Working with Data
## 7. Read in titanic dataset. Use str to see summary
```{r}
titanic<-read.csv(file='C:/Users/student/Documents/MATH421/data/titanic.csv', header=TRUE, sep=',')
str(titanic)
```
##8. Print first 10 rows in markdown
```{r}
library(kableExtra)
knitr::kable(titanic[1:10,], format='markdown')
```
##9. Count missing values in data, count missing values in each columns
```{r}
x<-is.na(titanic)
sum(x)
colSums(x)
```
##10. Average Age of passengers
```{r}
colMeans(titanic['Age'],na.rm=TRUE)
```
##11. Replace missing values by the mean
```{r}
titanicreplace<-replace(titanic, is.na(titanic),colMeans(titanic['Age'],na.rm=TRUE))
```
##12. Remove Name, ID, Ticket, Cabin
```{r}
newtitanic<-titanicreplace[-c(1,4,9,11)]
```
##13. Mean age of female passengers
```{r}
femtitanic<-subset(newtitanic, Sex=='female')
colMeans(femtitanic['Age'])
```
##14. Median fare of class 1 Passengers
```{r}
classone<-subset(newtitanic, Pclass==1)
classfare<-classone$Fare
median(classfare)
```
##15. Median fare of non Class 1 female passengers
```{r}
nonclassfemale<-subset(newtitanic, Pclass!=1 & Sex=='female')
nonclassfemalefare<-nonclassfemale$Fare
median(nonclassfemalefare)
```
##16. Calculate the median age of survived passengers who are female and Class 1 or Class 2
```{r}
survfemnon3<-subset(newtitanic,Pclass!=3 & Sex=='female' & Survived==1)
survfemnon3age<-survfemnon3$Age
median(survfemnon3age)
```
##17. Calculate the mean fare of female teenagers survived passengers
```{r}
femteensurv<-subset(newtitanic,Sex=='female' & Survived==1 & Age>=13 & Age <20)
colMeans(femteensurv['Fare'])
```
##18. Calculate the mean fare of female teenagers survived passengers for each class
```{r}
femteensurv<-subset(newtitanic,Sex=='female' & Survived==1 & Age>=13 & Age <20)
means<-aggregate(femteensurv['Fare'],list(femteensurv$Pclass),mean)
means
```
##19. . Calculate the ratio of Survived and not Survived for passengers who pays more than the average fare
```{r}
abovefare<-newtitanic[newtitanic['Fare']>mean(newtitanic$Fare,na.rm=TRUE),]
abovefare #shows there are 211 rows in the abovefare dataset
sum(abovefare['Survived'])/(211-sum(abovefare['Survived'])) #ratio of survived to not survived aka there are 1.48 people who survived to 1 person who did not survive
```
##20. . Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare
```{r}
newtitanic$sfare<-(newtitanic$Fare-mean(newtitanic$Fare,na.rm=TRUE))/sd(newtitanic$Fare,na.rm=TRUE)
```

##21. Add categorical variable named cfare that takes value cheap for passengers paying less the average fare and takes value expensive for passengers paying more than the average fare.
```{r}
newtitanic$cfare<-ifelse(newtitanic$Fare<mean(newtitanic$Fare,na.rm=TRUE),'Cheap','Expensive')
```
##22. Add categorical variable named cage that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on
```{r}
newtitanic$cage<-99
newtitanic$cage[newtitanic$Age<10]<-0
newtitanic$cage[newtitanic$Age>=10&newtitanic$Age<20]<-1
newtitanic$cage[newtitanic$Age>=20&newtitanic$Age<30]<-2
newtitanic$cage[newtitanic$Age>=30&newtitanic$Age<40]<-3
newtitanic$cage[newtitanic$Age>=40&newtitanic$Age<50]<-4
newtitanic$cage[newtitanic$Age>=50&newtitanic$Age<60]<-5
newtitanic$cage[newtitanic$Age>=60&newtitanic$Age<70]<-6
newtitanic$cage[newtitanic$Age>=70&newtitanic$Age<80]<-7
newtitanic$cage[newtitanic$Age>=80]<-8
newtitanic
```
##23. Show the frequency of Ports of Embarkation. It appears that there are two missing values in the Embarked variable. Assign the most frequent port to the missing ports. Hint: Use the levels function to modify the categories of categorical variables.
```{r}
table(newtitanic$Embarked)#Check most frequent port in the dataset
newtitanic$Embarked[newtitanic$Embarked=='']<-'S'
table(newtitanic$Embarked)
```
