---
title: "Math421_Assignment8"
author: "Louis Dion"
date: "10/28/2019"
output: html_document
---
#### input anything we need for the assignment
```{r}
library(purrr)
library(tidyverse)
library(readxl)
titanic = read_csv('C:/Users/student/Documents/MATH421/data/titanic.csv')
titanic$Pclass = as.character(titanic$Pclass)
titanic$Survived = as.character(titanic$Survived)
```


1. Working with the Titanic dataset. Use the map function to compute the number of NA for each columns. Use map_int to turn the output to an interger vector.
```{r}
na<-map(titanic, ~sum(is.na(.)))
navector<-map_int(titanic,~sum(is.na(.)))
navector
```

2. Use map function to calculate the mean of numeric variables in the Titanic dataset.
```{r}
average<-titanic%>%select_if(is.numeric)%>%map(., ~mean(.,na.rm=TRUE))
average
```

3. Use map function to show the means, medians, and standard deviations of numeric variables in the Titanic dataset. Use map_df to show the result in a data frame format.
```{r}
df<-titanic%>%select_if(is.numeric)%>%map(., ~list(
  mean =  mean(.x,na.rm=TRUE),
  median = median(.x,na.rm=TRUE),
  sd = sd(.x,na.rm=TRUE)))
df1<-titanic%>%select_if(is.numeric)%>%map_df(.,~list(
  mean =  mean(.x,na.rm=TRUE),
  median = median(.x,na.rm=TRUE),
  sd = sd(.x,na.rm=TRUE))) 
df1$names<-colnames(titanic%>%select_if(is.numeric))

df1
```

4. Use map function to show the mode (most frequent category) and the number of categories for each categorical variables. Put the result in data frame.
```{r}
#function to calculate mode in a vector
Mode = function(x){
    ta = table(x)
    tam = max(ta)
    if (all(ta == tam))
         mod = NA
    else
         if(is.numeric(x))
    mod = as.numeric(names(ta)[ta == tam])
    else
         mod = names(ta)[ta == tam]
    return(mod)
}

#using map
cat<-titanic%>%select_if(is.character)%>%map_df(.,~list(
  mode = Mode(.x)[1],
  categories = length(unique(.x))))
cat$names<-colnames(titanic%>%select_if(is.character))

cat

```

5. Use map to graph the histograms of numeric variables and bar charts of categorical variables in the Titanic data
```{r}
map(titanic%>%select_if(is.numeric),  ~hist(.))
#X NAMES
```
```{r}
map(titanic%>%select_if(is.character),  ~barplot(table(.)))
#X NAMES
```

6. Redo Question 11 in Assignment 7 using map. Hint: You may want to use a nested map. For instance this following codes print out all the possibility of x+y where both x and y are running from 1 to 2.

map(1:2, function(x){map(1:2, function(y){x+y})})
```{r}
redo<- function(d,k) {
  if (k==1) {
    df1 = d[, sapply(d, function(col) length(unique(col))) <= 4]
    map(1:ncol(df1), function(i) {map(1:ncol(df1),function(j) {print(names(df1)[c(i,j)])
      print(ggplot(df1,aes(x = df1[[i]], fill = df1[[j]])) + geom_bar(position = 'dodge')+labs(x = names(df1)[[i]], fill = names(df1)[[j]]))})})
  }
  if (k==2) {
    df1 = d[, sapply(d, function(col) is.numeric(col))]  
    map(1:ncol(df1), function(i) { print (names(df1)[i])
      print (ggplot(df1, aes(x = df1[[i]]))+geom_density()+labs(x = names(df1)[[i]]))})
  }
  if (k==3) {
    df1 = d[, sapply(d, function(col) is.numeric(col))]
    df2 = d[, sapply(d, function(col) length(unique(col))) <= 4]
    map(1:ncol(df1), function(i) {map(1:ncol(df2), function(j) {print(ggplot(data=df1,aes(x = df1[[i]]))+geom_density(data = df2, aes(col = df2[[j]]))+labs(x = names(df1)[[i]], col = names(df2)[[j]]))})})
  }
}
redo(titanic,3)
```


7. The function lm computes a linear model. For instance, to do run linear model on the output (y variable) Fare of the input Age (x variable) in the titanic data frame, d, we use
# To run a linear model of Fare on Age
lm(Fare~Age, d)

# To get the r-square of the linear model
summary(lm(Fare~Age, d))$r.square
Write a function, using map, to

input: a data frame
output: a data frame of consist three columns x1, x2 and r where the pairs (x1,x2) are all possible pairs of continuous variables of the input data frame and r is the corresponding r square of x1 and x2. Test the function on the titanic data assuming that the PassengerID is also a coninuous variable.

```{r}

seven<- function(d){
d1 = d%>%select_if(is.numeric)
d2 = data.frame() 
  for (i in 1:ncol(d1)) {
    for (j in 1:ncol(d1)) {
      d3 = map_df(d1,~list(
        x1=names(d1)[i],
        x2=names(d1)[j],
        r = round(summary(lm(d1[[j]]~d1[[i]],d1))$r.square,digits = 5)
      ))
      d3 = d3[1,]
      d2=rbind(d2,d3)
  }
} 
  return (d2)
}

seven(titanic)



```

